Solve LMI with saturation 1 dimensional case

### 2024-10-23 Journal/notes.tex
Progress on Journal

With the ETM mechanism described as follow:

@@

@@

@@

>>>>

With the ETM mechanism described as follows:

@@

And $\eta$ dynamic is defined as:

$$
  \eta^{+} = (\rho + \lambda) \eta - \psi(\omega_m, \chi)
$$

This formulation ensures $\eta$ to be always positive and made it possible to consider it as a positive system like in \cite{positive-systems}. 
As a quick proof I can say that if the triggering condition $\psi$ is negative then the positiveness of $\eta$ is trivial. If instead $\psi$ is positive and $\psi \geq \rho \eta $

@@

It works by putting $\alpha = 1$

@@

It works by putting $\alpha = 1$

\subsection*{Attempt on handcrafting NN}
Considerations on it can be found in the last subsection of \textit{"Linear integrator 1 dimension case"}.
I am now trying to train a 3-layer NN with deep learning method on a K matrix found with LQR method and not pole placement. It will then be attempted to:
\begin{itemize}
  \item Find a LMI solution for the system with integrator and non-linearity for the 3 layer network. If it doesn't work simplify the problem and add complexity bit by bit
  \item Test the K gain from LQR method to non-linear system with and without saturation
  \item Find the LMI solution for the non-linear system with input saturation: $u = sat(K_{LQR} \cdot x)$
\end{itemize}

\subsection*{1 layer saturation controller to non-linear pendulum with integrator in the state}
The system under analysis is represented by the dynamic:

$$
m l^{2} \ddot{\theta} = m g l\ sin(\theta) - \mu \dot{\theta} + u
$$

That discretized becomes:

$$
  \begin{cases}
    \theta^{+} = \theta + \dot{\theta} \delta t\\
    \dot{\theta}^{+} = \dot{\theta} + \ddot{\theta} \delta t = \dot{\theta} + (\frac{g}{l} sin(\theta) - \frac{\mu}{m l^{2}} \dot{\theta}) \delta t + \frac{dt}{m l^{2} u} \\
    \eta^{+} = \eta + y = \eta + (\theta - r)
  \end{cases}
$$

I want to deal with the non-linearity in the form of $sin(\theta) - \theta$. In order to do so, I add and subtract the term $\frac{g}{l} \cdot dt \cdot \theta$ to $\dot{\theta}$ dynamics. Then by expressing $\phi(\theta) = sin(\theta) - \theta$ I can rewrite the overall dynamics as:

\begin{equation}
  \bmx{
    \theta^{+} \\
    \dot{\theta}^{+} \\
    \eta^{+}
  } = \bmx{
    1 & \delta t & 0 \\
    \frac{g \cdot dt}{l} & 1 - \frac{\mu \cdot dt}{m l^{2}} & 0 \\
    1 & 0 & 1
  } \bmx{
    \theta \\
    \dot{\theta} \\
    \eta
  } + \bmx{
    0 \\
    \frac{dt}{m l^{2}} \\
    0
  } u + \bmx{
    0 \\
    \frac{g \cdot dt}{l} \\
    0 
  } \phi 
\end{equation}

With more condensed notation:
$$
  x^{+} = A x + B u + C \phi
$$

The sector conditions on the saturation are the same of \cite{css-extended} adapted for the single saturation case:
$$
N = \bmx{
\begin{array}{c|c|c}
  0 & 1 & 0 \\
  \hline
  K & 0 & 0
\end{array}
}
$$

Each element is null except for $N_{u \omega} = 1$ and $N_{\nu x} = K$ so to have:
\begin{align*}
  \nu &= N_{\nu x} \cdot x + N_{\nu \omega} \cdot \omega + N_{\nu b} = K x\\
  u &= N_{ux} \cdot x + N_{u \omega} \cdot \omega + N_{u b} = \omega = sat(\nu)
\end{align*}

In the end the control input is $u = sat(\nu) = sat(K x)$

I consider a quadratic Lyapunov function $V = \tilde{x}\tr P \tilde{x}$ and I want to bound $\Delta V$ to be strictly negative.

### 2024-10-23 Linear integrator 1 dimension case/ideas.tex
Progress on Journal

>>>>

We have explored some way to craft it. The problem lies in the dimensionality of the matrices, given the feasibility of the 1 dimensional case:

$$
  u = sat(Kx)
$$

We wanted to extend the problem to an $l$ layer controller by designing something like:

$$
  u = sat_1 \left( W_1 \left( sat_0 \left( W_0 x \right) \right) \right)
$$

Such that, staying in the linear zone of the saturation, we have $u = W_1 W_0 x = K x$ and to design something like $W_1 = k W_0^{-1}$ with $W_0$ and $W_1$ to be defined. Furthermore, we have to ensure that after the first layer we keep staying inside the linearity region. To do so we have to force something like this:

$$
  -\bar{\nu}_i \leq W_i \nu \leq \bar{\nu}_i
$$

For each layer, this directly translates to the inclusion of vector $\nu$ to the following set: \\ $ S_i = \left\{ \nu \in \RR^{n_{\phi}} : \nu\tr W_i\tr W_i \nu < \bar{\nu}^{2} \right\}$
Then we can express the ellipsoid conditions:

$$
  \nu\tr \frac{W_i\tr W_i}{\bar{\nu}^{2}} \nu \leq 1
$$

Since we want the inclusion of one linearity zone into the following we impose for every layer:

$$
  \nu\tr \frac{W_{i-1}\tr W_{i-1}}{\bar{\nu}_{i-1}^{2}} \nu \geq \nu\tr \frac{W_{i}\tr W_{i}}{\bar{\nu}_{i}^{2}} \to W_{i-1}\tr W_{i-1} - W_{i}\tr W_{i} \frac{\bar{\nu}_{i-1}^{2}}{\bar{\nu}_{i}^{2}} \geq 0
$$

Applying Schur's complement:

$$
  \bmx{
    W_{i-1}\tr W_{i-1} & W_{i}\tr \\
    W_{i} & \frac{\bar{\nu}_{i}^{2}}{\bar{\nu}_{i-1}^{2}} \Id
  } \geq 0
$$

Everything seems to be fine by fixing $\bar{\nu}_0$ equal to the actuator saturation level, but the problem lies in the dimensionality of the matrices $W_i$ since we need their inverse to compute the matrix for the following layer and to be invertible they need to be square. This enforces a strong limit on the architecture of the network since it forces to have the dimension of the input, the state, equal to the number of each hidden layer and also to the dimension of the output. Hence, this method is not applicable to the general case of a multi-layer network controller.
